
Generally SHYFEM is run with a fixed time step given by the
parameter |idt|.
This choice is acceptable when the model runs in unconditionally
stable conditions (ie. linear simulation, no horizontal viscosity).

The introduction of the advective terms (|ilin=0|) or horizontal
viscosity (|ahpar| greater 0)
can introduce instabilities. To be sure that the model runs in stable 
conditions, it must be assured that the Courant Number is smaller than 1. 
Please note that only in the case of advection we should call
this number the Courant number. However, we will continue to use
the term Courant number for all stability related issues.

In the case of advection the Courant number is defined as
 \begin{equation}
 Cou=\frac{v\Delta t}{\Delta x}
 \end{equation}
where $v$ is the current speed, $\Delta t$ the time step and $\Delta x$
the element size. For finite elements, due to the triangular grid, this
expression is slightly more complicated. As can be seen, lowering the time step
will bring the Courant number below the limit of 1.

To keep the Courant Number under the limit it is necessary to adapt the 
time step at every computation. The variable timestep is computed
introducing in the |STR| file in the |$para| section the parameters
|itsplt|, |coumax| and |idtsyn|.

|coumax| gives the limit of the Courant number. This is normally 1, but since
no exact stability limit can be derived for the non-linear advective terms,
another value can be specified. If instabilities arise, a slightly lower
value than 1 (0.9) can be tried.

|itsplt| decides about the time step splitting.
If this value
               is 0, the time step will be kept constant at its initial
               value. A value of 1 devides the initial time step into
               (possibly) equal parts, but makes sure that at the end
               of the micro time steps one complete macro time
               step has been executed. The last mode |itsplt| = 2
               does not care about the macro time step, but always
               uses the biggest time step possible. In this case
               it is not assured that after some micro time steps
               a macro time step will be recovered. Please note
               that the initial macro time step |idt| will never be exceeded.

Finally, the parameter |idtsyn| is only used in case of |itsplt| = 2. 
This parameter makes sure that
               after a time of |idtsyn| the time step will be syncronized
               to this time. Therefore, setting |idtsyn| = 3600 means
               that there will be a time stamp every hour, even if the model
               has to take one very small time step in order to reach that
               time.

An example of how to set the variable time stepping scheme is shown
in \Fig\figref{vartime}. Here the Courant number is lowered to 0.9 and
the variable time step is syncronized every 3600 seconds (1 hour).

\begin{figure}[ht]
\begin{verbatim}
$para
        coumax = 0.9   itsplt = 2   idtsyn = 3600
$end
\end{verbatim}
\caption{Example of variable time step settings. The time step is syncronized
at every hour, and the Courant number is lowered to 0.9.}
\label{fig:vartime}
\end{figure}

